Scalar_M1

### 2015-05-22 models/Model_Playspace.Rmd
Possible transformations for priors, MSE for alpha, additional plotting. Some code re-arranged ordering of data-load and model run.
Messing around with addition of negative valence, adjusting valence, scalar sd and mean - need to figure out why knitr isn't working

```{r}
rm(list=ls())
source("analysis/useful_dplyr.R")
library(rjson)
```

@@

Read Data 
---------

```{r}
priors = read.csv("models/prior.csv")
# Possible transformations
#priors$prior.p = sqrt(priors$prior.p)/(sum(sqrt(priors$prior.p)))
priors$prior.p = (1/(-log(priors$prior.p))) / sum((1/(-log(priors$prior.p))))

speaker = read.csv("models/speaker.csv")
speaker.alt = read.csv("models/e5a.csv")
listener = read.csv("models/listener.csv")
```

@@

  if (useprior) {
    prior <- priors[rating, "prior.p"]
  } else { #unif prior
    prior = .2 
  }

  speaker.prob(rating, degree, m, alpha) * prior
}

@@

  if (usenone) {
    mat$none <- c(.8, .2, 0, 0, 0)
    mat$bad <- c(.25, .5, .25, 0, 0)
  }

@@

```{r}
# Which alpha has tightest correlation with human judgements
alphas = seq(from=1, to=20)
fit = sapply(alphas, FUN=function(n) {
  md <- data %>%
    do(run.rsa(., alpha = n, useprior = FALSE, usenone = TRUE))

  mean((md$pred - md$listener.p)^2)
})
qplot(alphas, fit, main="MSE for alpha values", ylab="MSE")
best.alpha = which(fit == min(fit))
best.alpha

# RUN MODEL
md <- data %>%
  do(run.rsa(., alpha = best.alpha, useprior = FALSE, usenone = TRUE))
```

@@

```{r}
# Facets all scales, Model vs Human judgements
qplot(stars, listener.p, col=degree, 
      data = md, main="Model vs Human judgments") + 
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 2)

@@

Priors
```{r}
qplot(c(1:5), priors$prior.p,
      geom="bar", stat = "identity",
      binwidth = 1, main="Priors over ratings")
```

>>>>

```{r}
rm(list=ls())
setwd("~/Desktop/Projects/scalar_implicature/")
source("analysis/useful_dplyr.R")
library(rjson)
```

@@

Read Data ---->
---------

```{r}
priors = read.csv("models/prior.csv")
# Possible transformations --->
priors$prior.p = sqrt(priors$prior.p)/(sum(sqrt(priors$prior.p)))
#priors$prior.p = (1/(-log(priors$prior.p))) / sum((1/(-log(priors$prior.p))))
#priors$prior.p = (log(priors$prior.p + 1)) / sum((log(priors$prior.p + 1)))
speaker = read.csv("models/speaker.csv")
#speaker.alt = read.csv("models/e5a.csv")
listener = read.csv("models/listener.csv")
```

@@

  if (useprior) {
    prior <- priors[rating, "prior.p"]
  } else { #unif prior
    prior = .2 
  }
  speaker.prob(rating, degree, m, alpha) * prior
}

@@

  if (usenone) {
    mat$none <- c(.9, .1, 0, 0, 0)
    mat$neg_valence <- c(.2, .5, .2, .1, 0)
  }

@@

```{r}
# Which alpha has tightest correlation with human judgements
alphas = seq(from=1, to=20)
fit = sapply(alphas, FUN=function(n) {
  md <- data %>%
    do(run.rsa(., alpha = n, useprior = TRUE, usenone = TRUE))

  mean((md$pred - md$listener.p)^2)
  #cor(md$pred, md$listener.p)
})
qplot(alphas, fit, main="MSE for alpha values", ylab="MSE")
best.alpha = which(fit == min(fit))
best.alpha

# RUN MODEL
md <- data %>%
  do(run.rsa(., alpha = best.alpha, useprior = TRUE, usenone = TRUE))
```

@@

```{r}
# Facets all scales, Model vs Human judgements
qplot(stars, listener.p, col=degree, 
      data=md, main="Model vs Human judgments",
      ylab="Posterior judgments p(r|u)") + 
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

@@

Speaker informativity measures
```{r}
# Scalar judgment scores by rating
qplot(x = stars, y = speaker.judgment, col = degree, ymax = 5,
      geom = "line", stat = "identity", 
      position = "dodge", 
      data = speaker, main= "Scalar + Rating compabililty") + 
  geom_linerange(aes(ymin = speaker.judgment - speaker.cil, 
                     ymax = speaker.judgment + speaker.cih)) + 
  facet_wrap(~scale)

# Scalar compatibility (information ratings)
qplot(x = stars, y = speaker.p, col = degree, ymax = 1,
      geom = c("point", "smooth"), stat = "identity", 
      position = "dodge", 
      data = speaker, main= "Scalar Compaibility distr") + 
  facet_wrap(~scale)
```

Pragmatic Listener Judgments
```{r}
qplot(x=stars, y=listener.p, col=degree, ymax=1,
      geom=c("point", "smooth"), stat = "identity",
      ylab="Compatibility", position="dodge", 
      data=listener, main="Pragmatic listener judgements") + 
  facet_wrap(~scale)
```

Model judgments
```{r}
qplot(x = stars, y = pred, col = degree, ymax = 1, ylab="Posterior judgments",
      geom = c("point", "smooth"), stat = "identity", 
      position = "dodge", 
      data = md, main= "Scalar + Rating compabililty") + 
  facet_wrap(~scale)
```

Priors
```{r}
qplot(c(1:5), priors$prior.p,
      geom="bar", stat = "identity",
      binwidth = 1, main="Priors over ratings")

shapiro.test(priors$prior.p)
ks.test(priors$prior.p, "pnorm", mean=mean(priors$prior.p), sd=sd(priors$prior.p))
```

```{r}
#Some properties of our scales
scales = levels(d$scale)[1:5]
scales.info = sapply(scales, FUN=function(s) {
    curr <- data %>%
      filter(scale == s, degree == "hi") %>%
      select(speaker.p)
    return( c(sd(as.numeric(unlist(curr[,2]))),
              mean(as.numeric(unlist(curr[,2])))))
})
row.names(scales.info) = c("sd", "mean")
scales.info <- t(scales.info)
```
